---
title: "project1C5"
author: "sourabh jain"
date: "04/06/2020"
output: html_document
---

# Project 1

## Intructions for the assignment
1.Code for reading in the dataset and/or processing the data   
2.Histogram of the total number of steps taken each day    
3.Mean and median number of steps taken each day     
4.Time series plot of the average number of steps taken   
5.The 5-minute interval that, on average, contains the maximum number of steps  
6.Code to describe and show a strategy for imputing missing data   
7.Histogram of the total number of steps taken each day after missing values are imputed     
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends     
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report      


### 1.Code for reading in the dataset and/or processing the data
```{r readingcode}
setwd("C:/Users/Srb/Documents/Sourabh")
activity<-read.csv("activity.csv")
```

#### Exploring data
```{r explore}
dim(activity)
str(activity)
names(activity)
head(activity)
sum(is.na(activity))
# with lubridate package change the date formate
library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))
```

### 2. Histogram of the total number of steps taken each day

```{r plot1}
total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(total_steps) <- c("date", "steps")
hist(total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

### 3. Mean and median number of steps taken each day 
```{r MeanAndMedian}
mean(total_steps$steps)
median(total_steps$steps)
```

### 4.Time series plot of the average number of steps taken 
```{r timeSeries}
average_activity<-with(activity, aggregate(steps, by=list(interval), FUN= mean,na.rm=TRUE))
names(average_activity)<-c("interval","avg_step")
plot(average_activity$interval,average_activity$avg_step, type="l", xlab="Interval", ylab="Average number of step", lwd=2, main="time series plot")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps 

```{r}
average_activity[which.max(average_activity$avg_step),1]
```


### 6.  Code to describe and show a strategy for imputing missing data 
### 7.  Histogram of the total number of steps taken each day after missing values are imputed

I decided to fill in all of the missing values in the dataset by the mean number of steps per interval.

```{r}
m<-mean(average_activity$avg_step)
missingIndex<-is.na(activity[,1])
activity1<-activity
activity1[missingIndex,1]<-m
total_steps_imputed<-with(activity1, aggregate(steps, by=list(date), FUN=sum))
names(total_steps_imputed)<-c("date","step")
hist(total_steps_imputed$step, xlab="Steps per day", ylab="count",main = "histogram with imputed missing value", col="yellow", breaks = seq(0,25000,by=1000))
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends 

```{r}
activity1$date<-as.Date(activity1$date)
library(dplyr)
activity2<-activity1%>%
        mutate(dayType= ifelse(weekdays(activity1$date)=="Saturday" | weekdays(activity1$date)=="Sunday", "Weekend", "Weekday"))
head(activity2)

averageStepByDayTypeAndInterval<-activity2 %>%
  group_by(dayType, interval) %>%
  summarize(averageStepByDay=sum(steps))

head(averageStepByDayTypeAndInterval)

library(lattice)
with(averageStepByDayTypeAndInterval, 
      xyplot(averageStepByDay ~ interval | dayType, 
      type = "l",      
      main = "Total Number of Steps within Intervals by dayType",
      xlab = "Daily Intervals",
      ylab = "Average Number of Steps"))
'''




